---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



# Installation

Install from Github.

```{r eval=FALSE}
remotes::install_github("AlexWeinreb/rshannonca")
```

## Note on version

If you have an older version of scipy (fails in 1.5, works in 1.7), you will encounter this error:

```
Error in py_call_impl(callable, dots$args, dots$keywords) : 
  TypeError: rankdata() got an unexpected keyword argument 'axis'
```

That is the case if you have an older version of Python (fails with 3.6, works with 3.8). See `reticulate`'s documentation for details on how to solve it, the following should work:

```{r eval=FALSE}
reticulate::install_python(version = "3.9.6")
reticulate::use_python_version("3.9.6", required = TRUE)
reticulate::virtualenv_create(envname = "shannonca", packages = "shannonca", version = "3.9.6")
```

In that case, you can then load your installed environment before calling `RunSCA()` with:

```{r eval=FALSE}
reticulate::use_virtualenv("shannonca")
```


# Usage

Interacts directly with a Seurat object, adding an object in the `reductions` slot (just like `RunPCA()` or `RunUMAP()`).

```{r}
library(Seurat)
library(rshannonca)


reticulate::use_virtualenv("shannonca") #or adapt to your local Python environment

data("pbmc_small")
pbmc_small <- NormalizeData(pbmc_small)
pbmc_small <- FindVariableFeatures(pbmc_small)
pbmc_small <- ScaleData(pbmc_small)

```


Now perform the SCA

```{r}

pbmc_small <- RunSCA(pbmc_small)

DimPlot(pbmc_small, reduction = "sca")

```

We can directly cluster on the results of the SCA:

```{r}
pbmc_small <- FindNeighbors(pbmc_small, reduction = "sca")
pbmc_small <- FindClusters(pbmc_small)
```

Note that we decomposed the SCA by default into 50 dimensions (or the number of genes if it is smaller), we can visualize more than 2.

```{r}

DimPlot(pbmc_small, dims = 1:2, reduction = "sca")

DimPlot(pbmc_small, dims = c(2,3), reduction = "sca")
```


```{r eval=FALSE}
# in 3D, cannot be displayed on Github
plot.data <- FetchData(object = pbmc_small,
                       vars = "seurat_clusters") |>
  cbind(Embeddings(object = pbmc_small,
                   reduction = "sca"))

plotly::plot_ly(data = plot.data, 
                x = ~SC_1, y = ~SC_2, z = ~SC_3, 
                color = ~seurat_clusters,
                type = "scatter3d", 
                mode = "markers", 
                marker = list(size = 5, width=2), # controls size of points
                text=~seurat_clusters, #This is that extra column we made earlier for which we will use for cell ID
                hoverinfo="text") #When you visualize your plotly object, hovering your mouse pointer over a point shows cell names


rm(plot.data)
```

And we can compare clustering based on SCA and the standard one based on PCA:

```{r}
# save SCA clusters
pbmc_small$sca_clusters <- pbmc_small$seurat_clusters

# compute PCA clusters
pbmc_small <- RunPCA(pbmc_small, verbose = FALSE)
pbmc_small <- FindNeighbors(pbmc_small, reduction = "pca", verbose = FALSE)
pbmc_small <- FindClusters(pbmc_small, verbose = FALSE)
DimPlot(pbmc_small, reduction = "pca")
pbmc_small$pca_clusters <- pbmc_small$seurat_clusters


# Compare
library(ggplot2)
library(dplyr)

pbmc_small@meta.data |>
  group_by(sca_clusters, pca_clusters) |>
  summarize(nb_cells = n(), .groups = "drop") |>
  ggplot() +
  geom_point(aes(x = sca_clusters, y = pca_clusters, size = nb_cells)) +
  scale_x_discrete(guide = guide_axis(angle = 90))
```

